---
title: "Practical Machine Learning Project"
---

# Introduction

The course project data is provided by:

> Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. [Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements](http://groupware.les.inf.puc-rio.br/work.jsf?p1=10335). Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 

[Source:Human Activity Recognition](http://groupware.les.inf.puc-rio.br/har)


# Loading required libraries
```{r, warning=FALSE, error=FALSE, echo=FALSE}
library(AppliedPredictiveModeling)
library(caret)
library(rattle)
library(rpart.plot)
library(ggplot2)
library(randomForest)
library(e1071)
```

#Input Data
Downloaded the data from the source and loaded for analysis.
```{r}
rm(list = ls(all = TRUE))

trainingRaw <- read.csv(file="pml-training.csv", header=TRUE, as.is = TRUE, stringsAsFactors = FALSE, sep=',', na.strings=c('NA','','#DIV/0!'))
testingRaw <- read.csv(file="pml-testing.csv", header=TRUE, as.is = TRUE, stringsAsFactors = FALSE, sep=',', na.strings=c('NA','','#DIV/0!'))

trainingRaw$classe <- as.factor(trainingRaw$classe)  
```

#Cleaning data
Removing data not needed for the analysis especially null or 0 values.
```{}
NAindex <- apply(trainingRaw,2,function(x) {sum(is.na(x))}) 
trainingRaw <- trainingRaw[,which(NAindex == 0)]
NAindex <- apply(testingRaw,2,function(x) {sum(is.na(x))}) 
testingRaw <- testingRaw[,which(NAindex == 0)]
```

#Processing variables
```{r}

v <- which(lapply(trainingRaw, class) %in% "numeric")

preObj <-preProcess(trainingRaw[,v],method=c('knnImpute', 'center', 'scale'))
trainLess1 <- predict(preObj, trainingRaw[,v])
trainLess1$classe <- trainingRaw$classe

testLess1 <-predict(preObj,testingRaw[,v])

#Remove non-zero values
nzv <- nearZeroVar(trainLess1,saveMetrics=TRUE)
trainLess1 <- trainLess1[,nzv$nzv==FALSE]

nzv <- nearZeroVar(testLess1,saveMetrics=TRUE)
testLess1 <- testLess1[,nzv$nzv==FALSE]
```

#Create cross validation set
```{r}
#The training set is divided in two parts, one for training and the other #for cross validation

set.seed(12031987)

inTrain = createDataPartition(trainLess1$classe, p = 3/4, list=FALSE)
training = trainLess1[inTrain,]
crossValidation = trainLess1[-inTrain,]
```

#Train model
```{r}
#Train model with random forest due to its highly accuracy rate. The model is build on a training set #of 28 variables from the initial 160. Cross validation is used as train control method.

modFit <- train(classe ~., method="rf", data=training, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE )

#Accuracy on training set and cross validation set

#Following the computation on the accuracy of trainig and cross validation set

#Training set:

trainingPred <- predict(modFit, training)
confusionMatrix(trainingPred, training$classe)
#Confusion Matrix and Statistics

cvPred <- predict(modFit, crossValidation)
confusionMatrix(cvPred, crossValidation$classe)


```

#RESULT. Answer to the questions.
```{r}
 #Predictions on the real testing set

testingPred <- predict(modFit, testLess1)
testingPred
```


